Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2016-01-17 cordillera-cycle.tex
Typeset: replace equation environments by align.

      A~pseudo-plastic sliding law,
\begin{equation}
    \label{eqn:pseudoplastic}
    \vec{\tau}_{\mathrm{b}} = -\tau_{\mathrm{c}} \frac{\vec{v}_{\mathrm{b}}}{{v_{\text{th}}}^q\,|\vec{v}_{\mathrm{b}}|^{1-q}} \,,
\end{equation}
      relates the bed-parallel shear stresses, $\vec{\tau}_{\mathrm{b}}$, to the
      sliding velocity, $\vec{v}_{\mathrm{b}}$ (Table~\ref{tab:params}). The yield
      stress, $\tau_{\mathrm{c}}$, is modelled using the Mohr--Coulomb criterion,
\begin{equation}
   \tau_{\mathrm{c}} = c_0 + N\,\tan{\phi} \,,
\end{equation}
      where cohesion, $c_0$, is assumed to be zero. The friction angle,
      $\phi$, varies from 15 to 45\degree as a~piecewise-linear function of
      modern bed elevation, with the lowest value occuring below modern sea
      level (0\,\unit{m} above sea level, \unit{m\,a.s.l.}) and the highest
      value occuring above the generalised elevation of the highest
      shorelines \citep[200\,\unit{m\,a.s.l.},][Fig.~5]{Clague.1981}, thus
      accounting for a~weakening of till associated with the presence of
      marine sediments (cf. \citealp{Martin.etal.2011};
      \citealp[supplement]{Aschwanden.etal.2013};
      \citealp{PISM-authors.2015}). Effective pressure, $N$, is related to
      the ice overburden stress, $\rho gh$, and the modelled amount of
      subglacial water, using a~formula derived from laboratory experiments
      with till extracted from the base of Ice Stream B in West Antarctica
      \citep{Tulaczyk.etal.2000, Bueler.Pelt.2015},
\begin{equation}
    N = \delta \rho gh \, 10^{(e_0/C_{\mathrm{c}}) (1 - (W/W_{\text{max}}))} \,,
\end{equation}
      where $\delta$ is chosen as 0.02, $e_0$ is a~measured reference void
      ratio and $C_{\mathrm{c}}$ is a~measured compressibility coefficient
      (Table~\ref{tab:params}). The amount of water at the base, $W$, varies
      from zero to $W_{\text{max}}=2$\,\unit{m}, a~threshold above which additional
      melt water is assumed to drain off instantaneously. Finally, the
      bedrock topography responds to ice load following a~bedrock
      deformation model that includes local isostasy, elastic lithosphere
      flexure and viscous astenosphere deformation in an infinite half-space
      \citep{Lingle.Clark.1985,Bueler.etal.2007}. A~relatively low viscosity
      value of $\nu_{m} = 1\times10^{19}$\,\unit{Pa\,s} is used for the
      astenosphere (Table~\ref{tab:params}) in accordance with the results
      from regional glacial isostatic adjustment modelling at the northern
      Cascadia subduction zone \citep{James.etal.2009}.

@@

      The PDD computation accounts for stochastic temperature variations by
      assuming a~normal temperature distribution of standard deviation
      $\sigma$ aroung the expected value $T_{\mathrm{m}}$. It is expressed by an
      error-function formulation \citep{Calov.Greve.2005},
\begin{equation}
    \label{eqn:calovgreve}
    {\text{PDD}} = \int_{t_1}^{t_2} \mathrm{d}t
        \left[\frac{\sigma}{\sqrt{2\pi}}
                \exp\left({-\frac{T_{\mathrm{m}}^2}{2\sigma^2}}\right)
              + \frac{T_{\mathrm{m}}}{2} \, {\text{erfc}}
                \left(-\frac{T_{\mathrm{m}}}{\sqrt{2}\sigma}\right)\right] \,,
\end{equation}
      which is numerically approximated using week-long sub-intervals. In
      order to account for the effects of spatial and seasonal variations of
      temperature variability \citep{Seguinot.2013}, $\sigma$ is computed
      from NARR daily temperature values from 1979 to 2000
      \citep{Mesinger.etal.2006}, including variability associated with the
      seasonal cycle (Fig.~\ref{fig:atm}). Degree-day factors for snow and
      ice melt are derived from mass-balance measurements on contemporary
      glaciers from the Coast Mountains and Rocky Mountains in British
      Columbia \citep[Table~\ref{tab:params};][]{Shea.etal.2009}.

@@

      Climate forcing driving ice sheet simulations consists of
      a~present-day monthly climatology,
$\{T_{\mathrm{m}0}, P_{\mathrm{m}0}\}$, where temperatures are modified by offset time series,
${\Delta}T_{\text{TS}}$, and lapse-rate corrections, ${\Delta}T_{\text{LR}}$:
\begin{align}
    T_{\mathrm{m}}(t, x, y) &= T_{\mathrm{m}0}(x, y) + {\Delta}T_{\text{TS}}(t)
                    + {\Delta}T_{\text{LR}}(t, x, y) \,, \\
    P_{\mathrm{m}}(t, x, y) &= P_{\mathrm{m}0}(x, y) \,.
\end{align}

@@

      Temperature offset time-series, ${\Delta}T_{\text{TS}}$, are derived from
      palaeo-temperature proxy records from the Greenland Ice Core Project
      \citep[GRIP,][]{Dansgaard.etal.1993}, the North Greenland Ice Core
      Project \citep[NGRIP,][]{Andersen.etal.2004}, the European Project for
      Ice Coring in Antarctica \citep[EPICA,][] {Jouzel.etal.2007}, the
      Vostok ice core \citep{Petit.etal.1999}, and Ocean Drilling Program
      (ODP) sites 1012 and 1020, both located off the coast of California
      \citep{Herbert.etal.2001}. Palaeo-temperature anomalies from the GRIP
      and NGRIP records were calculated from oxygen isotope
      (\chem{\delta^{18}O}) measurements using a~quadratic equation
      \citep{Johnsen.etal.1995},
\begin{align}
    {\Delta}T_{\text{TS}}(t) = &-11.88 [\chem{\delta^{18}O}(t)
                                -\chem{\delta^{18}O}(0)] \nonumber \\
                        &-0.1925[\chem{\delta^{18}O}(t)^2
                                 -\chem{\delta^{18}O}(0)^2] \,,
\end{align}
      while temperature reconstructions from Antarctic and oceanic cores
      were provided as such. All records were scaled linearly
      (Table~\ref{tab:records}) in order to simulate comparable ice extents
      at the LGM (Table~\ref{tab:extrema}) and realistic outlines
      (Fig.~\ref{fig:lr_maps}).

>>>>

      A~pseudo-plastic sliding law,
\begin{align}
&\label{eqn:pseudoplastic}
    \vec{\tau}_{\mathrm{b}} = -\tau_{\mathrm{c}} \frac{\vec{v}_{\mathrm{b}}}{{v_{\text{th}}}^q\,|\vec{v}_{\mathrm{b}}|^{1-q}} \,,
\end{align}
      relates the bed-parallel shear stresses, $\vec{\tau}_{\mathrm{b}}$, to the
      sliding velocity, $\vec{v}_{\mathrm{b}}$ (Table~\ref{tab:params}). The yield
      stress, $\tau_{\mathrm{c}}$, is modelled using the Mohr--Coulomb criterion,
\begin{align}
&\tau_{\mathrm{c}} = c_0 + N\,\tan{\phi} \,,
\end{align}
      where cohesion, $c_0$, is assumed to be zero. The friction angle,
      $\phi$, varies from 15 to 45\degree as a~piecewise-linear function of
      modern bed elevation, with the lowest value occuring below modern sea
      level (0\,\unit{m} above sea level, \unit{m\,a.s.l.}) and the highest
      value occuring above the generalised elevation of the highest
      shorelines \citep[200\,\unit{m\,a.s.l.},][Fig.~5]{Clague.1981}, thus
      accounting for a~weakening of till associated with the presence of
      marine sediments (cf. \citealp{Martin.etal.2011};
      \citealp[supplement]{Aschwanden.etal.2013};
      \citealp{PISM-authors.2015}). Effective pressure, $N$, is related to
      the ice overburden stress, $\rho gh$, and the modelled amount of
      subglacial water, using a~formula derived from laboratory experiments
      with till extracted from the base of Ice Stream B in West Antarctica
      \citep{Tulaczyk.etal.2000, Bueler.Pelt.2015},
\begin{align}
&N = \delta \rho gh \, 10^{(e_0/C_{\mathrm{c}}) (1 - (W/W_{\text{max}}))} \,,
\end{align}
      where $\delta$ is chosen as 0.02, $e_0$ is a~measured reference void
      ratio and $C_{\mathrm{c}}$ is a~measured compressibility coefficient
      (Table~\ref{tab:params}). The amount of water at the base, $W$, varies
      from zero to $W_{\text{max}}=2$\,\unit{m}, a~threshold above which additional
      melt water is assumed to drain off instantaneously. Finally, the
      bedrock topography responds to ice load following a~bedrock
      deformation model that includes local isostasy, elastic lithosphere
      flexure and viscous astenosphere deformation in an infinite half-space
      \citep{Lingle.Clark.1985,Bueler.etal.2007}. A~relatively low viscosity
      value of $\nu_{m} = 1\times10^{19}$\,\unit{Pa\,s} is used for the
      astenosphere (Table~\ref{tab:params}) in accordance with the results
      from regional glacial isostatic adjustment modelling at the northern
      Cascadia subduction zone \citep{James.etal.2009}.

@@

      The PDD computation accounts for stochastic temperature variations by
      assuming a~normal temperature distribution of standard deviation
      $\sigma$ aroung the expected value $T_{\mathrm{m}}$. It is expressed by an
      error-function formulation \citep{Calov.Greve.2005},
\begin{align}
&\label{eqn:calovgreve}
    {\text{PDD}} = \int_{t_1}^{t_2} \mathrm{d}t
        \left[\frac{\sigma}{\sqrt{2\pi}}
                \exp\left({-\frac{T_{\mathrm{m}}^2}{2\sigma^2}}\right)
              + \frac{T_{\mathrm{m}}}{2} \, {\text{erfc}}
                \left(-\frac{T_{\mathrm{m}}}{\sqrt{2}\sigma}\right)\right] \,,
\end{align}
      which is numerically approximated using week-long sub-intervals. In
      order to account for the effects of spatial and seasonal variations of
      temperature variability \citep{Seguinot.2013}, $\sigma$ is computed
      from NARR daily temperature values from 1979 to 2000
      \citep{Mesinger.etal.2006}, including variability associated with the
      seasonal cycle (Fig.~\ref{fig:atm}). Degree-day factors for snow and
      ice melt are derived from mass-balance measurements on contemporary
      glaciers from the Coast Mountains and Rocky Mountains in British
      Columbia \citep[Table~\ref{tab:params};][]{Shea.etal.2009}.

@@

      Climate forcing driving ice sheet simulations consists of
      a~present-day monthly climatology,
$\{T_{\mathrm{m}0}, P_{\mathrm{m}0}\}$, where temperatures are modified by offset time series,
${\Delta}T_{\text{TS}}$, and lapse-rate corrections, ${\Delta}T_{\text{LR}}$:
\begin{align}
&T_{\mathrm{m}}(t, x, y) = T_{\mathrm{m}0}(x, y) + {\Delta}T_{\text{TS}}(t)
                    + {\Delta}T_{\text{LR}}(t, x, y) \,, \\
&    P_{\mathrm{m}}(t, x, y) = P_{\mathrm{m}0}(x, y) \,.
\end{align}

@@

      Temperature offset time-series, ${\Delta}T_{\text{TS}}$, are derived from
      palaeo-temperature proxy records from the Greenland Ice Core Project
      \citep[GRIP,][]{Dansgaard.etal.1993}, the North Greenland Ice Core
      Project \citep[NGRIP,][]{Andersen.etal.2004}, the European Project for
      Ice Coring in Antarctica \citep[EPICA,][] {Jouzel.etal.2007}, the
      Vostok ice core \citep{Petit.etal.1999}, and Ocean Drilling Program
      (ODP) sites 1012 and 1020, both located off the coast of California
      \citep{Herbert.etal.2001}. Palaeo-temperature anomalies from the GRIP
      and NGRIP records were calculated from oxygen isotope
      (\chem{\delta^{18}O}) measurements using a~quadratic equation
      \citep{Johnsen.etal.1995},
\begin{align}
{\Delta}T_{\text{TS}}(t) ={~}&-11.88 [\chem{\delta^{18}O}(t)
                                -\chem{\delta^{18}O}(0)] \nonumber \\
                        &-0.1925[\chem{\delta^{18}O}(t)^2
                                 -\chem{\delta^{18}O}(0)^2] \,,
\end{align}
      while temperature reconstructions from Antarctic and oceanic cores
      were provided as such. All records were scaled linearly
      (Table~\ref{tab:records}) in order to simulate comparable ice extents
      at the LGM (Table~\ref{tab:extrema}) and realistic outlines
      (Fig.~\ref{fig:lr_maps}).
